Limited evidence for widespread heterogeneity in psychology

### 2018-06-27 writing/draft-heterogeneity.rmd
Changed tables to simple versions, addied figure captions
To arrange in word later

---
title: "Limited evidence for widespread heterogeneity in psychology"
author: "Anton Ohlsson Collentine"
header-includes:
   - \setlength\parindent{24pt}
   - \usepackage{setspace}
   - \doublespacing
output: pdf_document
---

@@

```{r load data and packages}
library(readr)
library(dplyr)
library(kableExtra)
library(metafor)
library(purrr)

dat <- read_csv("../data/collated_summary_data.csv")
```

```{r prep-heterogeneity-table}
#library(metafor)
#library(dplyr)
#library(purrr)

est_heterogen_smd_raw <- function(x){
  if(any(x[, "effect_type"] == "Risk difference")){ #without the 'any' we will get warnings because we apply 'if' to a vector

    fit <- rma(measure = "RD", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2,  n1i = ntreatment, n2i = ncontrol,  data = x)

  } else if(any(x[, "outcomes1_2"] == "mean _ SE")){

    fit <-  rma(yi = effect_size, sei = outcome_c2,  data = x)

  } else if(any(x[, "effect_type"] == "Raw mean difference")){
    if(any(x[, "rs"] %in% c("RRR5", "RRR7"))){ #For these us the Knapp and Hartung adjustment of standard errors

      fit <- rma(measure = "MD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol, test ="knha", data = x)
    } else{
      fit <- rma(measure = "MD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol, data = x)
    }

  } else if(any(x[, "outcomes1_2"] == "mean _ SD")){

    fit <- rma(measure = "SMD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol, data = x)

  } else if(any(x[, "effect_type"] == "r")){
    if(any(x[, "rs"] == "ML1")){
      fit <- rma(measure = "COR", ri = effect_size, ni = Ntotal,  data = x)
    } else{ #for ML3
      fit <- rma(measure = "UCOR", ri = effect_size, ni = Ntotal,  data = x, vtype = "UB")
      #Note that ML3 specifies measure = "COR" with vtype = "UB", but that this is specified as above in the latest metafor 
    }

  } else{ #for all two-group count effects

    fit <- rma(measure = "OR2DL", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2, 
               n1i = ntreatment, n2i = ncontrol,  data = x)
    #standardized, ML1 used 'OR2D' which = 'OR2DL' in latest version of metafor

  }

  I2_1 <- confint(fit)$random[3, ] #Gives us the I2 estimate and its confidence interval
  data.frame(eff_size = fit$b[[1]], #effect size (point estimate)
             s_I2 = I2_1[1], s_ci.lb = I2_1[2], s_ci.ub = I2_1[3]) #I2 + CI
}

#Apply function to data
res <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") #apply function, rbind results into dataframe ('dfr'), make sure purrr is up to date

OR <- c("Allowed vs. forbidden", "Gain vs. loss framing", "Norm of reciprocity", "Low vs. high category scales") #Effects reported as d in many labs 1 but meta-analyzed as OR

#Add rs to the data-frame
het <- dat %>% 
  select(rs, Site, effect, effect_type) %>% 
  group_by(rs, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  select(rs, effect,k , everything()) %>% 
  arrange(desc(s_I2),  desc(s_ci.ub)) #sort by I2 and upper CI bound
```
Empirical research is typically portrayed as proceeding in two stages. First, belief in the existence of an effect is established. Second, its generalizability is examined by exploring its boundary conditions. In the first stage, inferential statistics are used  to minimize the risk that a discovery is due to sampling error.  In the second stage, one may ask to what extent the effect depends on a particular choice of four contextual factors; the 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables [ref: Campbell & Stanley]. This extent is often explored through replications of the original study that are either as similar as possible to the original (called 'direct' or 'exact' replications) or with some deliberate variation on conceptual factors (so-called 'conceptual' or 'indirect' replications) [ref: Zwaan et al], and once sufficient studies have accumulated through meta-analysis. In meta-analysis, the heterogeneity of an effect size (henceforth referred to as heterogeneity) is a measure of an effect's susceptibility to changes in these four factors. An effect strongly dependent on one or more of the four factors, unless controlled for, should exhibit high heterogeneity. In this paper we examine what degree of heterogeneity can be expected in replication studies in psychology and explore a proposed relationship with effect size. We focus in this paper on direct replications.

@@

##Table 1.
###Heterogeneity may increase with effect size
```{r effect-size-heterogeneity-table}
#library(kableExtra)
rooted_relabilities <- matrix(rep(c(.8, .7,.6), 3), ncol = 3)
rho <- c(0, .5, .8)
rel <- c("R = .8", "R = .7", "R = .6")

rooted_relabilities <- sweep(rooted_relabilities, 2, rho, "*")
diff_row <- data.frame(rel = "Diff.", X1 = 0, X2 = 0.05, X3 = 0.08)

het_demo <- data.frame(rel, rooted_relabilities)
het_demo <- rbind(het_demo, diff_row)

het_demo %>% knitr::kable("latex", booktabs = T, col.names = c("", "Rho = 0", "Rho = .5", "Rho = .8")) %>% 
  kable_styling() %>% 
  footnote(threeparttable = TRUE,
           general = "R = square root of the multiplied reliabilities") %>% 
  row_spec(3, hline_after = T)

@@

We examine the evidence for widespread heterogeneity in psychology and explore the correlation between effect size and heterogeneity, using 10 pre-registered multi-lab replication projects in psychology (Table 2). These 10 projects examined a total of 37 primary outcome variables and arguably represent the best, least biased, meta-analytic data currently available in psychology. To better interpret heterogeneity estimates we also estimate power of each project to find zero/small/medium/large heterogeneity as defined by [Higgins et al]. Of the four contextual factors a study might vary on (sample population, settings,  treatment variables, and measurement variables) the studies in these projects vary only on two: sample population and settings. Nonetheless, if we believe an effect is sensitive to changes in these two factors we may still expect to find some heterogeneity. For example, the variety of countries and samples involved in some of the replication projects, not all of which are WEIRD (Western, Educated, Industrial, Rich, Democratic [ref: Weird]) might be a reason to expect heterogeneity. On the other hand, it may be that the less disparate collection of studies included in the these direct replication meta-analyses as compared to most psychological meta-analyses result in low heterogeneity estimates, similar to estimates in medicine. Regardless of outcome, the sampled multi-lab replication projects can only inform us of susceptibility to changes in the varied contextual factors  (sample population and settings), with no information on susceptibility to changes in other contextual factors.

##Table 2.
```{r data-summary-table}
#library(dplyr)
#library(kableExtra)

#Shortened version of APA reference
papers <- c('Klein, R. A., Ratliff, K. A., Vianello, M., Adams, R. B., Jr., Bahník, S., Bernstein, M. J., . . . Nosek, B. A. (2014). Investigating variation in replicability: A "many labs" replication project.', 
'Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., ... & Brown, E. R. (2016). Many Labs 3: Evaluating participant pool quality across the academic semester via replication.',
'Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., ... & Brown, E. R. (2016). Many Labs 3: Evaluating participant pool quality across the academic semester via replication.',
'Alogna, V. K., Attaya, M. K., Aucoin, P., Bahník, S., Birch, S., Birt, A. R., ... & Buswell, K. (2014). Registered replication report: Schooler and engstler-schooler (1990).',
'Eerland, A., Sherrill, A. M., Magliano, J. P., Zwaan, R. A., Arnal, J. D., Aucoin, P., ... & Crocker, C. (2016). Registered replication report: Hart & Albarracín (2011).',
'Hagger, M. S., Chatzisarantis, N. L., Alberts, H., Anggono, C. O., Batailler, C., Birt, A. R., ... & Calvillo, D. P. (2016). A multilab preregistered replication of the ego-depletion effect.',
'Cheung, I., Campbell, L., LeBel, E. P., Ackerman, R. A., Aykutoglu, B., Bahník, S., ... & Carcedo, R. J. (2016). Registered Replication Report: Study 1 from Finkel, Rusbult, Kumashiro, & Hannon (2002).',
'Wagenmakers, E. J., Beek, T., Dijkhoff, L., Gronau, Q. F., Acosta, A., Adams Jr, R. B., ... & Bulnes, L. C. (2016). Registered Replication Report: Strack, Martin, & Stepper (1988).',
'Bouwmeester, S., Verkoeijen, P. P., Aczel, B., Barbosa, F., Bègue, L., Brañas-Garza, P., ... & Evans, A. M. (2017). Registered Replication Report: Rand, Greene, and Nowak (2012).',
"O'Donnell, M., Nelson, L., McLatchie, N. M., & Lynott, D. J. (2017). Registered Replication Report: Dijksterhuis & van Knippenberg (1998)")

summary.table <- dat %>% 
  group_by(rs) %>% 
  summarize(k = n_distinct(Site),
            Countries = n_distinct(country), 
            Effects = n_distinct(effect),
            Participants = if(Effects == 1){sum(Ntotal)}
                           else{sum(Ntotal) / Effects}) %>% #average across effects
  mutate(Paper = papers) %>% 
  select(RS = rs, Paper, everything()) #change order of columns

summary.table %>% knitr::kable("latex", booktabs = T, digits = 0) %>% 
  kable_styling() %>% 
  column_spec(2, width = "25em") %>% #specify width of paper column
  footnote(threeparttable = TRUE,
           general = "For studies with several effects the number of participants is the average across effects, rounded to the closest whole number. Participant numbers are those used for primary analyses by original authors (i.e., after exclusions). RS = Replication Study, k = no. primary studies, ML = Many Labs, RRR = Registered Replication Report. (OSF.io/XXXX)")
```

###Method

We downloaded and collated summary data from 10 pre-registered multi-lab replication projects in psychology (Table 2). Data from all 10 projects were publicly available on the Open Science Framework (osf.io) and downloaded between 2018/02/01 and 2018/03/31. Although some projects (e.g. RRR4) reported results from several outcome variables, we only included primary outcome variables as stated in accompanying publications, resulting in a total of 37 effects. For each effect we extracted (osf.io/XXXX/) summary data (e.g., means and standard deviations) at the level of the lab as specified by the original authors for their primary analysis (i.e., typically after exclusions). We extracted information on the country of each lab, whether participants were physically in the lab for the study, total number of participants per lab, type of effect size, and additional information related to each effect (see codebook; osf.io/XXXX/). Extracted data were in a variety of formats: Excel (Many labs 1, RRR1 & RRR2), CSV (Many labs 3, RRR3, RRR4, RRR5, RRR6) and as PDF tables (RRR7, RRR8). In two cases (RRR5 and RRR6) it was necessary to download the raw data to extract summary data. Although a particular lab may have participated in several projects, the lab indicator was typically not the same across projects. Even so, we kept the original lab indicators to facilitate comparing observations in our dataset with the original datasets. Finally, we collated the summary data for all effects into one dataset for analysis (osf.io/XXXX/).

To examine heterogeneity across effects, we computed meta-analytic estimates for all 37 effects in our dataset (Table 3). We ran all analyses as specified by the replication authors (osf.io/XXXX/). All effects were estimated with random-effects models and the Restricted Maximum Likelihood (REML) estimator using the R-package metafor [ref], though with a variety of outcome variables: correlations, standardized mean differences, raw mean differences, odds ratios and risk differences. ML1 transformed odd ratios into standardized mean differences when meta-analyzing under the assumption that responses followed logistic distributions [ref: Sanchez-Meca; metafor]. Two projects (RRR5 and RRR7) used the Knapp and Hartung adjustment of the standard errors [ref: Knapp & Hartung] and ML3 correlations were corrected for bias [ref: Hedges 1989; metafor]. After estimating effects, ML3 transformed correlations into eta-squared for reporting, which we did not. For each effect we estimated I2 and its 95% confidence interval. Confidence intervals were calculated in metafor which uses the Q-profile method [ref Jackson 2014].

[Ignore the following paragraph - power simulation paragraph - subject to change]
To facilitate interpreting results, we estimated power and type 1 error for each of the 37 effects under zero/small/medium/large heterogeneity (I2 = 0/25/50/75%  respectively), as well as the expected I2-distributions across effects at these heterogeneity levels (osf.io/XXXX/). To do so we simulated results for each effect given the number of studies, k, and the sample sizes of each study, $N_k$, for that effect. Our approach consisted of (i) varying the between-studies variance,  $\tau^2$, for each effect to find the values that best corresponded to I2-estimates of 25/50/75% and (ii) using the found values to estimate power and type 1 error by computing the proportion non-significant results over a large number of replications. In step (i), we varied $\tau^2$ between 0.000 and 0.250 in increments of 0.005,  and replicated each condition 50 times.  We then computed the average I2 across the 50 replications, discarding non-converging models, and for each effect saved the $\tau^2$ values that best corresponded to small, medium, and large I2.  For homogeneity we set $\tau^2$ to zero. In step (ii), to estimate power and type 1 error, we ran one thousand replications at each identified level of heterogeneity using the $\tau^2$-values of step (i). We considered a result significant when the lower bound of the 95% I2 confidence interval was above zero and save heterogeneity estimates were saved from each replication for plotting I2-distributions.

[Ignore the following paragraph - power simulation paragraph - subject to change]
In both steps of our simulation, we followed the same data generating procedure. For each effect, we generated k effect sizes, $\theta_k$, from a normal distribution with a mean of zero and between-studies variance $\tau^2$: $\theta_k \sim N(0,  \tau^2)$.  For the sake of simplicity, we made the choice to set the average effect size to zero and treated each effect as consisting of a treatment and control group. We thus drew an effect size $\theta_k$ for each replication study of an effect, and within each study we then generated $N_k / 2$ (rounded to closest integer) sample values for a treatment and for a control group from two normal distributions with means $\mu_t = \theta_k$ and $\mu_c = 0$ and equal within-study variances ($\sigma^2 = 1$). This provided sample means and standard deviations for all k treatment and control groups of an effect, which we then used to estimate a random effects model for standardized mean differences in metafor using REML. This process was repeated the desired number of times for each condition and effect.

To examine the correlation between effect size and I2 we converted all effect sizes to a common metric (osf.io/XXXX/). Since I2 is logically and mathematically bounded at zero, we also correlated effect sizes with the closely related heterogeneity estimate H2, which is calculated as $H2 = Q  / (k - 1)$ and equals $1 / (1 - I2)$ for $I2 \geq 0$. We first converted all effects into correlations (r) and used the R-package metafor [ref] to estimate I2 and meta-analytic effects expressed as Fisher's z. In doing so, we fitted random-effects models with metafor's default REML estimator. Fisher's z is a normalizing correlation transformation that ranges from negative infinity to positive infinity, though except for extreme correlations it stays close to the -1 to 1 range [ref]. For mean differences we calculated the pooled standard deviation, [ref Borenstein in Cooper & Hedges, 2009, p. 226), standardized the effect size and converted it to a correlation (p. 234) with a correction factor for unequal sample sizes (p. 234). In one case (RRR8), we first had to convert reported standard errors into pooled standard deviation (p.224). For risk differences and odds ratios we first added 1/2 to a cell if it was empty to avoid dividing by zero, next calculated the logarithmic odds ratio (p. 266), converted this to Cohen's d (p. 232) and then finally to a correlation. All formulas used are presented in Appendix A.

Table 3 presents the meta-analytic estimates and I2 with confidence intervals for each of the 37 included effects.

##Table 3.
```{r heterogeneity-table}
#library(kableExtra)

het %>% knitr::kable("latex", align = "c", booktabs = TRUE,digits = 2, col.names = c("RS", "Effect", "k", "Effect type", "Estimate", "I2(%)", "ci.lb", "ci.ub")) %>% 
  kable_styling("scale_down") %>%
  footnote(threeparttable = TRUE, escape = FALSE,
           general = " Effects were estimated in metafor using REML. The following effects are odds ratios transformed into standardized mean differences: 'Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales'. RS = Replication Study, k = no. primary studies, Estimate = Point estimates of effect sizes, ci.lb = lower bound of I2 95\\\\% confidence interval, ci.ub = upper bound of I2 95\\\\% confidence interval. SMD = Standardized Mean difference (Hedge's g), MD = Mean Difference, RD = Risk Difference, r = correlation. (OSF.io/XXXX)")

We examined the evidence for heterogeneity across a set of pre-registered multi-lab replication projects. Few effects show evidence of heterogeneity. The lower bound I2 confidence interval excludes zero for only 7/37 effects (19%; Table 3), all part of the ML1 project. The percentage of heterogeneity estimates larger than 0 (23/37; 62%) suggests heterogeneity for at least some effects, as this percentage is higher than the expected frequency of non-zero estimates under homogeneity (46%, or about 17/37), based on the chi-square distribution and average k across projects. Of those effects whose lower bound confidence interval includes zero, 15/30 (50%) have confidence intervals that exceed 50% (medium heterogeneity).

[add segment on power with figures when ready]

Larger effect sizes appear to be associated with higher heterogeneity estimates. In Table 3, the four effects with highest I2 estimates also have the largest effect sizes. As expected, heterogeneity tends to increase with (absolute) effect size, with effect size sufficient to explain 61% of variation in I2 estimates according to a simple regression model (Figure 1). The closely related, but unbounded, heterogeneity estimate H2 provides a similar result (R2 = .53). Excluding Anchoring effects as robustness check makes little difference (R2 = .51 for I2).

##Figure 1
```{r}
knitr::include_graphics("../code/effect_het.png")
```

@@

There are some additional limits to the generalizability of claims based on the data in our study. For one, the included effects are neither a representative nor random sample of effects in psychology and as such do not support making strong claims about average heterogeneity levels in psychology. In respect to this point we look forward to the forthcoming Many Labs 2 project which should provide considerable new data on the prevalence of heterogeneity in psychology. In addition, while both the I2 confidence intervals and distribution of point estimates support a small number of heterogenous effects amongst those studied, most confidence intervals are congruent with anything from low to high heterogeneity as defined by [Higgins et al]. As always, absence of evidence is not evidence of absence. Finally, since the included studies only vary on two contextual factors (sample population and settings) , the data tell us nothing about the included effects' susceptibility to changes in the other two factors (treatment and measurement variables).

>>>>

---
title: "Limited evidence for widespread heterogeneity in psychology"
author: "Anton Ohlsson Collentine"
header-includes:
   - \setlength\parindent{24pt}
   - \usepackage{setspace}
   - \doublespacing
output: word_document
---

@@

```{r load packages, data and source functions}
if(!require(readr)){install.packages("readr")}
if(!require(metafor)){install.packages("metafor")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(purrr)){install.packages("purrr")}
if(!require(kableExtra)){install.packages("kableExtra")}

library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data iteration
library(metafor) #To run meta-analyses
library(kableExtra) #For creating good-looking pdf tables

dat <- read_csv("../data/collated_summary_data.csv")

source("../code/helper_functions_tables_figures.r") #Load functions to prep data for figures
```
Empirical research is typically portrayed as proceeding in two stages. First, belief in the existence of an effect is established. Second, its generalizability is examined by exploring its boundary conditions. In the first stage, inferential statistics are used  to minimize the risk that a discovery is due to sampling error.  In the second stage, one may ask to what extent the effect depends on a particular choice of four contextual factors; the 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables [ref: Campbell & Stanley]. This extent is often explored through replications of the original study that are either as similar as possible to the original (called 'direct' or 'exact' replications) or with some deliberate variation on conceptual factors (so-called 'conceptual' or 'indirect' replications) [ref: Zwaan et al], and once sufficient studies have accumulated through meta-analysis. In meta-analysis, the heterogeneity of an effect size (henceforth referred to as heterogeneity) is a measure of an effect's susceptibility to changes in these four factors. An effect strongly dependent on one or more of the four factors, unless controlled for, should exhibit high heterogeneity. In this paper we examine what degree of heterogeneity can be expected in replication studies in psychology and explore a proposed relationship with effect size. We focus in this paper on direct replications.

@@

Table 1.

_Heterogeneity may increase with effect size_

We examine the evidence for widespread heterogeneity in psychology and explore the correlation between effect size and heterogeneity, using 10 pre-registered multi-lab replication projects in psychology (Table 2). These 10 projects examined a total of 37 primary outcome variables and arguably represent the best, least biased, meta-analytic data currently available in psychology. To better interpret heterogeneity estimates we also estimate power of each project to find zero/small/medium/large heterogeneity as defined by [Higgins et al]. Of the four contextual factors a study might vary on (sample population, settings,  treatment variables, and measurement variables) the studies in these projects vary only on two: sample population and settings. Nonetheless, if we believe an effect is sensitive to changes in these two factors we may still expect to find some heterogeneity. For example, the variety of countries and samples involved in some of the replication projects, not all of which are WEIRD (Western, Educated, Industrial, Rich, Democratic [ref: Weird]) might be a reason to expect heterogeneity. On the other hand, it may be that the less disparate collection of studies included in these direct replication meta-analyses as compared to most psychological meta-analyses result in low heterogeneity estimates, similar to estimates in medicine. Regardless of outcome, the sampled multi-lab replication projects can only inform us of susceptibility to changes in the varied contextual factors  (sample population and settings), with no information on susceptibility to changes in other contextual factors.

Table 2.

_Pre-registered multi-lab replication studies_

##Method

We downloaded and collated summary data from ten pre-registered multi-lab replication projects in psychology (Table 2). Data from all ten projects were publicly available on the Open Science Framework (osf.io) and downloaded between 2018/02/01 and 2018/03/31. Although some projects (e.g. RRR4) reported results from several outcome variables, we only included primary outcome variables as stated in accompanying publications, resulting in a total of 37 effects. For each effect we extracted (osf.io/XXXX/) summary data (e.g., means and standard deviations) at the level of the lab as specified by the original authors for their primary analysis (i.e., typically after exclusions). We extracted information on the country of each lab, whether participants were physically in the lab for the study, total number of participants per lab, type of effect size, and additional information related to each effect (see codebook; osf.io/XXXX/). Extracted data were in a variety of formats: Excel (Many labs 1, RRR1 & RRR2), CSV (Many labs 3, RRR3, RRR4, RRR5, RRR6) and as PDF tables (RRR7, RRR8). In two cases (RRR5 and RRR6) it was necessary to download the raw data to extract summary data. Although a particular lab may have participated in several projects, the lab indicator was typically not the same across projects. Even so, we kept the original lab indicators to facilitate comparing observations in our dataset with the original datasets. Finally, we collated the summary data for all effects into one dataset for analysis (osf.io/XXXX/).

To examine heterogeneity across effects, we computed meta-analytic estimates for all 37 effects in our dataset (Table 3). We ran all analyses as specified by the replication authors (osf.io/XXXX/). All effects were estimated with random-effects models and the Restricted Maximum Likelihood (REML) estimator using the R-package metafor [ref], though with a variety of outcome variables: correlations, standardized mean differences, raw mean differences, odds ratios and risk differences. ML1 transformed odd ratios into standardized mean differences when meta-analyzing under the assumption that responses followed logistic distributions [ref: Sanchez-Meca; metafor]. Two projects (RRR5 and RRR7) used the Knapp and Hartung adjustment of the standard errors [ref: Knapp & Hartung] and ML3 correlations were corrected for bias [ref: Hedges 1989; metafor]. After estimating effects, ML3 transformed correlations into eta-squared for reporting, which we did not. For each effect we estimated $I^2$ and its 95% confidence interval. Confidence intervals were calculated in metafor which uses the Q-profile method [ref Jackson 2014].

In order to facilitate interpretation of our results, we estimated power and type 1 error for each of the 37 effects under zero/small/medium/large heterogeneity ($I^2$ = 0/25/50/75%  respectively). In addition, we plotted $I^2$-distributions across effects at these heterogeneity levels and compared with all observed $I^2$ estimates. To do so we simulated results for each effect given the number of studies, k, and the sample sizes, $N_k$, of those studies (osf.io/XXXX/). In the case of correlations we 1) generated a true correlation $\rho_k$ from a normal distribution with  between studies standard deviation $\tau$ and mean effect size zero, $\rho_k \sim N(0,  \tau)$, and 2) generated an observed correlation expressed as Fisher's z $rz_k$ from a normal distribution with mean $\rho_k$ and standard deviation $\sqrt{1 / N_k - 3}$, $rz_k \sim N(\rho_k, \sqrt{1 / N_k - 3})$. For mean differences we assumed equal within study group sizes $n_c = n_t$ ($= N_k$ rounded to closest integer) and standard deviation $\sigma_c = \sigma_t$. We then 1) generated study effect sizes $\mu_k$ from a normal distribution with mean zero and standard deviation $\tau$, $\mu_k \sim N(0, \tau)$, and 2) generated control and treatment means from the sampling distributions $\bar{x}_c \sim N(0, 1 / \sqrt{n_c})$ and $\bar{x}_t \sim N(\mu_k, 1 / \sqrt{n_t})$ respectively. Group variances were generated from their sampling distributions as well, in accordance with $s^2_c = \chi^2 (n_c - 1) / (n_c - 1)$ and  $s^2_t = \chi^2 (n_t - 1) / (n_t - 1)$. Due to the difficulties of data generation for odds ratios and risk differences we treated these as mean differences. For each effect, our simulations then proceeded in three steps: i) we varied the between studies standard deviation, $\tau$, between 0.000 and 0.50 in increments of 0.005 and replicated each condition 1000 times, each time estimating $I^2$ with a random effects model with REML, ii) we computed the average $I^2$ in each condition, discarding non-converged models, and iii) ran 10 000 replications at the $\tau$ values in step _ii_ that best corresponded to an $I^2$ estimate of 25% (small), 50% (medium) and 75% (large). For homogeneity we set $\tau$ to zero. We considered a result significant when the lower bound of the 95% $I^2$ confidence interval was above zero.

To examine the correlation between effect size and $I^2$ we converted all effect sizes to a common metric (osf.io/XXXX/). We first converted all effects into correlations (r) and used the R-package metafor [ref] to estimate $I^2$ and meta-analytic effects expressed as Fisher's z. In doing so, we fitted random-effects models with metafor's default REML estimator. Fisher's z is a normalizing correlation transformation that ranges from negative infinity to positive infinity, though except for extreme correlations it stays close to the -1 to 1 range [ref]. For mean differences we calculated the pooled standard deviation, [ref Borenstein in Cooper & Hedges, 2009, p. 226), standardized the effect size and converted it to a correlation (p. 234) with a correction factor for unequal sample sizes (p. 234). In one case (RRR8), we first had to convert reported standard errors into pooled standard deviation (p.224). For risk differences and odds ratios we first added 1/2 to a cell if it was empty to avoid dividing by zero, next calculated the logarithmic odds ratio (p. 266), converted this to Cohen's d (p. 232) and then finally to a correlation. All formulas used are presented in Appendix A (osf.io/XXXX/).  Since $I^2$ is set to zero for the majority of cases under homogeneity (i.e., truncated), we also correlated effect size with with the closely related heterogeneity estimate H2 [ref: Higgins 2002] as a sensitivity analysis. The H2 index is calculated as $H2 = Q  / (k - 1)$ and equals $1 / (1 - $I^2$)$ for $I^2 \geq 0$. It ranges from zero to positive infinity, where a larger score indicates more heterogeneity, a score of one indicates homogeneity, and a score between zero and one thus indicates 'excessive' homogeneity.

Table 3 presents the meta-analytic estimates and $I^2$ with confidence intervals for each of the 37 included effects.

```{r prep-heterogeneity-table}
# library(metafor)
# library(dplyr)
# library(purrr)

#fit a random effects model to all empirical data
#Each set of input datapoints is meta-analyzed as per the original replication project
#Function 'est_heterogen_smd_raw' is sourced
res <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") #apply function, rbind results into dataframe ('dfr'), make sure purrr is up to date

#Add rs and k to the data-frame and prep for table
het <- dat %>% 
  select(rs, Site, effect, effect_type) %>% 
  group_by(rs, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD"))

#Load simulation results (power/type 1 error for all effects)
dens <- readRDS("../code/temp_sim_results_2.RDS")
names(dens) <-  sort(unique(dat$effect)) #names were alphabetized upon splitting in simulation so must sort

#Summary of simulation results, needs to be combined with main part of table
I2_ci_lb <- dens %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(ci.lb > 0)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(zero = '1', small = '2', medium = '3', large = '4')

#Combine dataframes to create full table
het <-  het %>% 
  left_join(I2_ci_lb) %>% 
  select(rs, effect,k , everything()) %>% 
  arrange(desc(s_I2),  desc(s_ci.ub)) #sort by I2 and upper CI bound

@@

Table 3.

_Heterogeneity across primary effects in ten multi-lab replication projects._
```{r heterogeneity-table}
#library(kableExtra)

het %>% knitr::kable(digits = 2, col.names = c("RS", "Effect", "k", "Effect type", "Estimate", "I2(%)", "ci.lb", "ci.ub", "Zero", "Small", "Medium", "Large")) 
```

**Note:**  Effects were estimated in metafor using REML. The following effects are odds ratios transformed into standardized mean differences: 'Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales'. RP = Replication Project, k = no. primary studies, Estimate = Point estimates of effect sizes, ci.lb = lower bound of $I^2$ 95\\% confidence interval, ci.ub = upper bound of $I^2$ 95\\% confidence interval. SMD = Standardized Mean difference (Hedge's g), MD = Mean Difference, RD = Risk Difference, r = correlation. Under simulated heterogeneity, Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75\\%) respectively.  (OSF.io/XXXX)"

###$I^2$ estimates and confidence intervals

There is limited evidence for heterogeneity across the examined effects. Of the 37 effects,  4/37 (11%) have $I^2$ estimates that best correspond to large heterogeneity ($I^2$ = 75%), 4/37 (11%) to medium heterogeneity ($I^2$ = 50%), 9/37 (24%)  to small heterogeneity ($I^2$ = 25%) and 17/37 (54%) to zero heterogeneity ($I^2$ = 0%). However, despite relatively large k and total sample size for most projects (see Table 2), Table 3 shows wide confidence intervals for many effects: 10/37 (27%) effects have confidence intervals congruent with small to large heterogeneity and 15/37 (41%) with zero to medium.  The lower bound $I^2$ confidence interval excludes zero for only 7/37 effects (19%; Table 3), all part of the ML1 project. The percentage of heterogeneity estimates larger than 0 (23/37; 62%) suggests heterogeneity for at least some effects, as this percentage is higher than the expected frequency of non-zero estimates under homogeneity (46%, or about 17/37), based on the chi-square distribution and average k across projects.

###$I^2$ and power

We estimated $I^2$ for a large range of between studies standard deviations for each of the 37 effects (Figure 1). The $I^2$ index increases non-linearly with heterogeneity and is relatively more sensitive to lower between studies variances. This appears to be less so in the case of  mean differences than for correlations, which correspond to the more steeply inclined curves in Figure 1.

```{r}
knitr::include_graphics("../figures/tau-I2.png")
```

_Figure 1. Result of simulation relating $I^2$-values to between studies standard deviation. Each line represent one of 37 effects. The more steeply inclined lines belong to correlations and the others to mean differences._

Power and type 1 error estimates for zero/small/medium/large heterogeneity are shown for each effect in Table 3. In all cases the type 1 error is conservative, as compared to the expected 5% error rate. Power to detect small heterogeneity was low, ranging between 18% (Professor priming) and  45% (Flag priming), with an average of 28%. Power to detect medium heterogeneity was generally good, with an average of 79% power, but was only around 60% for several effects with low K. Power to detect strong heterogeneity was excellent across the board.

Figure 2 shows the distribution of $I^2$ at different heterogeneity levels and a histogram of the observed effects. The shortest bars in the histogram correspond to a single effect. The considerable overlap of the distributions illustrates why sufficient power can be difficult to achieve, and why confidence intervals for $I^2$ are often wide. In particular the spread of the distribution under small heterogeneity is illustrative of the low power to be expected under such circumstances. Given the densities of the distributions (height of the curves) the majority of observed effects are most likely to have either zero or small heterogeneity.  Only two effects appear to be of medium heterogeneity and three of large.

```{r}
knitr::include_graphics("../figures/density-I2.png")
```

_Figure 2. Simulated $I^2$ densities across 37 effects for zero, small, medium, and large heterogeneity according to the definitions of [Higgins], and a histogram of the observed $I^2$ estimates for the 37 effects. Each simulated density consists of approximately 1,480,000 point estimates._

###$I^2$ and effect sizes

Larger effect sizes appear to be associated with higher heterogeneity estimates. In Table 3, the four effects with highest $I^2$ estimates also have the largest effect sizes. As expected, heterogeneity tends to increase with (absolute) effect size, with effect size sufficient to explain 61% of variation in $I^2$ estimates according to a simple regression model (Figure 3). The closely related, but unbounded, heterogeneity estimate $H^2$ provides a similar result ($R^2$ = .53). Excluding Anchoring effects as robustness check makes little difference ($R^2$ = .51 for $I^2$).

```{r}
knitr::include_graphics("../figures/effect_het.png")
```

_Figure 3. The correlation between A) $I^2$ and effect size and B) $H^2$ and effect size for 37 effects from ten pre-registered multi-lab replication projects._

@@

There are some additional limits to the generalizability of claims based on the data in our study. For one, the included effects are neither a representative nor random sample of effects in psychology and as such do not support making strong claims about average heterogeneity levels in psychology. In respect to this point we look forward to the forthcoming Many Labs 2 project which should provide considerable new data on the prevalence of heterogeneity in psychology. In addition, while both the I2 confidence intervals and distribution of point estimates support a small number of heterogenous effects amongst those studied, most confidence intervals are congruent with anything from low to high heterogeneity as defined by [Higgins et al]. As always, absence of evidence is not evidence of absence. Finally, since the included studies only vary on two contextual factors (sample population and settings), the data tell us nothing about the included effects' susceptibility to changes in the other two factors (treatment and measurement variables).
